---
title: "CS 422 HW2"
author: "Ramesh Chandra Reddy Musilanna"
output:
  html_notebook:
    toc: yes
    toc_float: yes
  html_document:
    toc: yes
    df_print: paged
---

## 2.    Programming Problem
### 2.1
```{r}
library("ISLR")
set.seed(1122)
index <- sample(1:nrow(Auto), 0.95*dim(Auto)[1])
train.df <- Auto[index,]
test.df <- Auto[-index,]
#index
#train.df
```

### Part 2.1 - a
```{r}
linear_reg <- lm(mpg ~.-name,data = train.df)
linear_reg
```
#### Part 2.1 - a - i
<p>The attribute "name" is to identify a row uniquely which is already achieved by index of array.<p>

#### Part 2.1 - a - ii
```{r}
lsummary <- summary(linear_reg)
lsummary

paste("R-sq value is ", signif(lsummary$r.squared, 2))
paste("Adjusted R-sq value is ", signif(lsummary$adj.r.squared, 2))
paste("RSE is ", signif(sqrt(sum(linear_reg$residuals**2)/(length(linear_reg$residuals)-(1+(length(linear_reg$coefficients)-1)))), 2))
paste("RMSE is ", signif(sqrt(sum(linear_reg$residuals**2)/(length(linear_reg$residuals))), 2))



```
<p>Variance is described well in the model as R-squared value is 0.82 which is close to 1</p>


#### Part 2.1 - a - iii
```{r}
res <- resid(linear_reg)
# which =1 gives plot of Residuals vs Fitted values
plot(linear_reg,which = 1)

```

#### Part 2.1 - a - iv
```{r}
library(ggplot2)
hist_res<- qplot(linear_reg$residuals,geom = "histogram",bins = 10) +
         labs(title = "Histogram of Residuals",x = "Residuals")
plot(hist_res)
```

<p>The histogram is symmetric. The residuals are normally distributed.</p>

### Part 2.1 - b
#### Part 2.1 - b - i
```{r}
# By seeing the lowest pr-values, 'displacement', 'weight', and 'year' appears to be the 3 most significant predictor attributes.
linear_reg_3p <-  lm(mpg ~ displacement + weight + year, data = Auto)
#linear_reg_3p
```
#### Part 2.1 - b - ii
```{r}
lr_3p_summ <- summary(linear_reg_3p)
lr_3p_summ

paste("R-sq value is ", signif(lr_3p_summ$r.squared, 2))
paste("RSE is ", signif(sqrt(sum(lr_3p_summ$residuals**2)/(length(lr_3p_summ$residuals)-(1+(length(lr_3p_summ$coefficients)-1)))), 2))
paste("RMSE is ", signif(sqrt(sum(lr_3p_summ$residuals**2)/(length(lr_3p_summ$residuals))), 2))
```
<p>The model is good. Variance is described well in the model as R-squared value is 0.81 which is close to 1. The RSE and RMSE values are also quite close to 0, however RSE,RMSE are increased by 0.1 than previous.</p>

#### Part 2.1 - b - iii
```{r}
plot(linear_reg_3p,which = 1)
```

#### Part 2.1 - b - iv
```{r}
hist_res_1<- qplot(linear_reg_3p$residuals,geom = "histogram",bins = 10) +
         labs(title = "Histogram of Residuals",x = "Residuals")
plot(hist_res_1)
```
<p>The histogram does appear to be bell-shaped with little skewed in right. However, the histogram show the residuals to be normally distributed.</p>

#### Part 2.1 - b - v
<p>Models (a) and (b) show similarities in performance with respect to variance. The histogram of residual show that. However, Even with 3 predictors from the original dataset in model (b) managed to get an accurate model.</p>

### Part 2.1 - c
```{r}
test_3p.df = test.df[c('displacement', 'weight', 'year')]


mpg_predicted <- predict(linear_reg_3p, newdata = test_3p.df)
#mpg_predicted
linear_reg_3p.df <- data.frame(mpg_predicted, test.df[c('mpg')])
colnames(linear_reg_3p.df) <- c('Prediction', 'Response')
linear_reg_3p.df
```

### Part 2.1 - d
```{r}
#Using Confidence Interval
confidence_count.df <- predict(linear_reg_3p, test_3p.df, interval = "confidence", level = 0.95)
confidence_count.df <- data.frame(linear_reg_3p.df, confidence_count.df)


confidence_count.df <- subset(confidence_count.df, select = c(1,2,4,5))
colnames(confidence_count.df) <- c("Prediction", "Response", "Lower", "Upper")

#confidence_count.df

counting_fun <- function(x,y,z) {
  if ((x <= y) & (y <= z) ) {
    return(1)
  } else {
    return(0)
  }
}
Matches= mapply(counting_fun, confidence_count.df$Lower, confidence_count.df$Prediction, confidence_count.df$Upper)
confidence_interval.df <- cbind(confidence_count.df,Matches)
confidence_interval.df

```
```{r}

paste("Total observations correctly predicted: ",nrow(confidence_interval.df[confidence_interval.df$Matches == 1, ]))
```

### Part 2.1 - e
```{r}
#Using Prediction Interval
prediction_count.df <- predict(linear_reg_3p, test_3p.df, interval = "prediction", level = 0.95)
prediction_count.df <- data.frame(linear_reg_3p.df, prediction_count.df)


prediction_count.df <- subset(prediction_count.df, select = c(1,2,4,5))
colnames(prediction_count.df) <- c("Prediction", "Response", "Lower", "Upper")

#prediction_count.df
#using Counting_fun from 2.1.d
Matches= mapply(counting_fun, prediction_count.df$Lower, prediction_count.df$Prediction, prediction_count.df$Upper)
prediction_interval.df <- cbind(prediction_count.df,Matches)
prediction_interval.df
```
```{r}

paste("Total observations correctly predicted: ",nrow(prediction_interval.df[prediction_interval.df$Matches == 1, ]))
```
### Part 2.1 - f
#### Part 2.1 - f - i
<p>Number of matches using Confidence interval and Prediction Interval on test data is equal.</p>

#### Part 2.1 - f - ii
<p>As prediction interval is wider than the confidence interval, if mpg parameter in confidence interval matches  then mpg parameter in prediction interval also matches</p>